
##Analyzing Step Activity Measurements of Individuals

###Synopsis

Sign into GitHub using my credentials

fork https://GithUb.com/rdpeng/RepData_PeerAssessment1 to my account

https://github.com/NelsonBrown/RepData_PeerAssessment1

Create local repository for the project

mkdir "c\\Coursera\\Repdata_PeerAssessment1"

git remote https://github.com/NelsonBrown/RepData_PeerAssessment1

Open new file PA1_Template.Rmd in RStudio

Set working Directory to local repository directory

```{r}
setwd("C:/Coursera/RepData_PeerAssessment1")
```
###Loading and Processing the Raw Data


Download and Extract activity.zip to current directory creating activity.csv  

```{r}
setInternet2(TRUE)

fileUrl <- "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip?accessType=DOWNLOAD"
download.file(fileUrl,destfile = "\\activity.ZIP")
activity <- read.csv(unz("activity.zip", "activity.csv"))
```

Read activity.csv into activity data frame

```{r, echo=TRUE}
activity <- read.csv("activity.csv")
summary(activity)  
```

**As we can see from the summary there are 2304 step values missing**

Load ggplot2 library

```{r, echo=TRUE}
library(ggplot2)
```

Plot histogram of daily activity steps removing the missing data

```{r, echo=TRUE}
dailySteps <- aggregate(activity$steps, by=list(date=activity$date), FUN=sum, na.rm=TRUE)
colnames(dailySteps)[2] <- "steps"
qplot(steps, data = dailySteps, geom = "histogram", main = "Histogram of Daily Totals with missing data ignored", )
```

Calculate and plot Daily mean totals of steps removing the missing data
  
```{r, echo=TRUE}

dailyMeanSteps <- aggregate(activity$steps, by=list(date=activity$date), FUN=mean, na.rm=TRUE)

colnames(dailyMeanSteps)[2] <- "steps"


p <- qplot(date, steps, data = dailyMeanSteps, main = "Mean Daily Totals with missing data ignored", geom = c("point"), method = "lm", ylab="Step Total", xlab="Interval" )

p1 <- p + theme(
        axis.title.x = element_text(color="forestgreen", vjust=-0.35),
        axis.title.y = element_text(color="cadetblue" , vjust=0.35),   
        axis.text.x=element_text(angle=60, size=6, vjust=0.5)
)

p1
```

Calculate and plot Daily Median totals of steps removing the missing data
  
```{r, echo=TRUE}

dailyMedianSteps <- aggregate(activity$steps, by=list(date=activity$date), FUN=median, na.rm=TRUE)

colnames(dailyMedianSteps)[2] <- "steps"

p <- qplot(date, steps, data =dailyMedianSteps, main = "Median Daily Totals with missing data ignored", geom = "point", ylab="Steps", xlab="Interval")

p1 <- p + theme(
        axis.title.x = element_text(color="forestgreen", vjust=-0.35),
        axis.title.y = element_text(color="cadetblue" , vjust=0.35),   
        axis.text.x=element_text(angle=60, size=6, vjust=0.5)
)

p1
```

Calculate the average steps per interval period removing the missing data

```{r, echo=TRUE}

intervalSteps <- aggregate(activity$steps, by=list(interval=activity$interval), FUN=mean, na.rm=TRUE)

colnames(intervalSteps)[2] <- "steps"

 p <- qplot(interval, label = NULL, steps, data = intervalSteps, geom = "line", type = "1", main = "Daily Interval Totals with missing data ignored", ylab="Total Steps", xlab="By Interval")
 
 p1 <- p + theme(
          axis.ticks = element_blank(),
          axis.text.x = element_blank(),
          axis.title.x = element_text(color="forestgreen", vjust=-0.35),
          axis.title.y = element_text(color="cadetblue" , vjust=0.35),   
          axis.text.x=element_text(angle=60, size=6, vjust=0.5)
  )
 p1
```  

Fill in missing values with the mean steps value for a given interval  

```{r, echo=TRUE}

require(plyr)
require(Hmisc)

newactivity <- ddply(activity, "interval", mutate, steps = impute(steps, mean))

newactivity <- newactivity[order(newactivity$date,newactivity$interval),]
summary(newactivity)
```

Make the date value a factor

```{r, echo=TRUE}
newactivity$date <- as.factor(newactivity$date)
```

Plot histogram of daily activity steps with missing data filled in with generated data

```{r, echo=TRUE}
dailySteps <- aggregate(newactivity$steps, by=list(date=newactivity$date), FUN=sum)

colnames(dailySteps)[2] <- "steps"

qplot(steps, data = dailySteps, geom = "histogram", main = "Histogram of Daily Totals with No missing data", )
```

Calculate and plot mean daily totals of steps with the missing data filled in 

```{r, echo=TRUE}

dailyMeanSteps <- aggregate(activity$steps, by=list(date=activity$date), FUN=mean)

colnames(dailyMeanSteps)[2] <- "steps"


p <- qplot(date, steps, data = dailyMeanSteps, main = "Mean Daily Totals with no missing data", geom = "point", method = "lm", ylab="Daily Mean Steps", )

p1 <- p + theme(
        axis.title.x = element_text(color="forestgreen", vjust=-0.35),
        axis.title.y = element_text(color="cadetblue" , vjust=0.35),   
        axis.text.x=element_text(angle=60, size=6, vjust=0.5)
)

p1
````


Calculate and plot median daily totals of steps with the missing data filled in 

```{r, echo=TRUE}

 dailySMedianteps <- aggregate(newactivity$steps, by=list(date=newactivity$date), FUN=median)
 
 colnames(dailyMedianSteps)[2] <- "steps"
 
 p <- qplot(date, steps, data = dailyMedianSteps, main = "Median Daily Totals with no missing data", geom = "point", method = "lm", ylab="Daily Median Steps", )
 
 p1 <- p + theme(
         axis.title.x = element_text(color="forestgreen", vjust=-0.35),
         axis.title.y = element_text(color="cadetblue" , vjust=0.35),   
         axis.text.x=element_text(angle=60, size=6, vjust=0.5)
 )
 p1
```

Calculate interval steps mean with the missing data filled in

```{r, echo=TRUE}

intervalSteps <- aggregate(newactivity$steps, by=list(interval=newactivity$interval), FUN=mean)

colnames(intervalSteps)[2] <- "steps"

 p <- qplot(interval, label = NULL, steps, data = intervalSteps, geom + "line", type = "1", ylab="Steps", main = "Daily Interval Totals with no missing data", xlab="Interval", )
 
 p1 <- p + theme(
          axis.ticks = element_blank(),
          axis.text.x = element_blank(),
          axis.title.x = element_text(color="forestgreen", vjust=-0.35),
          axis.title.y = element_text(color="cadetblue" , vjust=0.35),   
          axis.text.x=element_text(angle=60, size=6, vjust=0.5)
  )
 p1
```

Add a column to the data frame specifying the day of the week
 Monday thru Sunday
 Convert Monday thru Friday to weekpart == "weekday"
 Convert Saturday and Sunday to "Weekend"

```{r, echo=TRUE}
newactivity$weekpart <- weekdays(as.Date(newactivity$date))
head(newactivity)
newactivity$weekpart[newactivity$weekpart=="Monday"] <- "weekday"
newactivity$weekpart[newactivity$weekpart=="Tuesday"] <- "weekday"
newactivity$weekpart[newactivity$weekpart=="Wednesday"] <- "weekday"
newactivity$weekpart[newactivity$weekpart=="Thursday"] <- "weekday"
newactivity$weekpart[newactivity$weekpart=="Friday"] <- "weekday"
newactivity$weekpart[newactivity$weekpart=="Saturday"] <- "weekend"
newactivity$weekpart[newactivity$weekpart=="Sunday"] <- "weekend"
newactivity$weekpart <- as.factor(newactivity$weekpart)
head(newactivity,10)
```

Plot the Weekend vs weekday activity

```{r, echo=TRUE}

par(mfrow = c(2,1),mfcol=c(2,1))

weekSteps <- aggregate(steps ~ interval + weekpart, newactivity, mean)

p <- qplot(interval, steps, data = weekSteps, ylab="Number of Steps", geom = "line", type ="l", facets = weekpart ~ .)
p1 <- p + theme(
                axis.title.x = element_text(color="forestgreen", vjust=-0.35),
                axis.title.y = element_text(color="cadetblue" , vjust=0.35),   
                axis.text.x=element_text(angle=50, size=8, vjust=0.5)
                )
p1
```  
load knitr library and invoke knit2html() to create md and html files 


 